---
title: GT tables
author: ~
date: '2019-11-08'
slug: post-about-gt-tables
categories: [R]
tags: [Rstats, gt tables, dplyr]
subtitle: ''
summary: ''
authors: []
lastmod: '2019-11-08T22:24:26-06:00'
featured: yes
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: [myblog]
draft: false
---

In my job, I make tables *all the time*  

I recently stumbled upon the {gt} package and whoa... these are some great looking tables! You can check out the vignettes [here](https://gt.rstudio.com/)

Even better, they are simple to append to any dplyr pipeline...

---

Step 1: Load {gt} and {dplyr}  
```{r echo = TRUE, message = FALSE, warning = FALSE}
library(gt)
library(dplyr)
```

&nbsp;  
Step 2:  

+ Create a simple frequency table, with mtcars, grouped by number of cylinders.  
+ Dplyr is fast and simple to use, but the console output is not so easy on the eyes...  
```{r echo = TRUE, message = FALSE, warning = FALSE}
mtcars %>%
  group_by(cyl) %>%
  summarize(N = n()) %>%
  mutate(Frequency = N/sum(N),
         Frequency = round(Frequency, 2))
```

&nbsp;  
Step 3: Lets add some {gt} magic...getting better  
```{r echo = TRUE, message = FALSE, warning = FALSE}
mtcars %>%
    group_by(cyl) %>%
    summarize(N = n()) %>%
    mutate(Frequency = N/sum(N),
           Frequency = round(Frequency, 2)) %>%
    gt() 
```

&nbsp;  
Step 4: Ok, now adding some bells and whistles, we can really make this table shine...
```{r}
mtcars %>%
  group_by(cyl) %>%
  summarize(N = n()) %>%
    mutate(Frequency = N/sum(N),
           Frequency = round(Frequency, 2)) %>%
    gt(rowname_col = "cyl") %>%
    tab_stubhead(label = "Cylinders") %>%
  tab_header(
    title = md("Table 1: MTCARS cyl table"),
    subtitle = md("Made with {gt} package")
  ) %>% 
  tab_footnote(
    footnote = "Added a footnote",
    locations = cells_data(
      columns = vars(N), 
      row = 1 #add to the row with 4 cylinders
    )
  ) %>% 
  tab_source_note(
    source_note = "Example 1: Add a source note here"
  ) %>%
  tab_source_note(
    source_note = "Example 2: Add another note/reference here!"
  ) %>%
  tab_options(heading.background.color = "#aa6f73", 
              heading.border.bottom.color = "black", 
              heading.title.font.size = 24, 
              heading.subtitle.font.size = 22, 
              table.width = 370, 
              table.font.size = 18, 
              column_labels.font.size = 20, 
              row.striping.include_stub = TRUE,
              row.striping.include_table_body = TRUE)
```

&nbsp;  
Done!  
Simply calling gt() creates the table, then add a few parameters to control the style/font size etc...Voila! You have a fancy gt Table...


